Categorical and gradient aspects of wordlikeness\thanks{Thanks to Gene Buckley, Constantine Lignos, Hilary Prichard, Charles Yang, and audiences at NELS 43 for helpful comments.}

### 2013-01-31 1.tex
Looks like this is wrapped up...!

Phonotactic knowledge is evidenced in quite different ways than phonological knowledge.
In Russian, for instance, a process of anticipatory assimilation ensures that derived clusters of obstruents agree in voice.\footnote{
    For sake of discussion, the complex behavior of [v] is ignored here.}

@@

\noindent
\citeauthor{PE} go on to label constraints on underlying representations as ``extensional'', and thus irrelevant to generative grammar, in contrast to ``intensional'' statements like phonological rules.
Regarding the facts about Russian obstruents discussed so far, it seems to be clear that the alternation facts are in some sense more privileged.
Were it the case that Russian speakers had not internalized a process of obstruent voicing assimilation, the only alternative explanation for the forms in (\ref{rovs}) is a massive system of phonologically conditioned suppletive allomorphy.
In contrast, were there no constraint on underlying obstruent voicing, however, it is not immediately obvious that anything would be different.
Presumably, this is what \citeauthor{PE} mean when they refer to constraints on underlying representations as ``computationally inert''.
There are also many cases where restrictions on underlying representations are specific to underlying representations and do not apply to surface representations.
For instance, hiatus is exceptionally rare in native Turkish roots; of the handful of examples in the Turkish Electronic Living Lexicon \citep{TELL}, many appear to be compounds (e.g., \emph{ısıalan} `endothermic', cf.~\emph{ısı} `heat', \emph{alan} `taker').
Yet numerous phonological processes give rise to hiatus in derived environments \citep[e.g.,][]{Kabak2007b}.
\citet{Silverman2000} observes that operations like reduplication and truncation also tend to introduce violations of restrictions that hold of underlying representations.
However, there are many facts which strongly suggest that speakers internalize phonotactic constraints.

@@

Speakers have difficulty producing \citep[e.g.,][]{Davidson2006a,Davidson2010,Rose2007,Vitevitch1997} and perceiving \citep{Dupoux1999,Kabak2007a,Massaro1983} certain types of nonce words judged to be phonotactically illicit.
For instance, \citet{GallagherInPress} finds that speakers of Quechua have difficulty repeating nonce words with multiple ejectives (e.g., [k'ap'i]), which do not occur in the language; in other words, multiple ejective sequences are not merely absent, but also difficult for these speakers to repeat.

In English, sequences of adjacent obstruents which do not also agree in voice (e.g., \emph{a}[b.s]\emph{inth}) are quite rare within a word, and therefore a hetero-voiced obstruent cluster is a clue to the presence of a word boundary in running speech.
Infants \citep[e.g.,][]{Mattys2001b} and adults \citep{McQueen1998b,Norris1997} are thought to use this heuristic for word recognition in experimental settings.\footnote{
    There is reason to distinguish the computations involved in using this word segmentation heuristic from those implicated by other phonotactic behaviors.
    Whatever the locus of [bnɪk]'s ill-formedness, for example, no segmentation into multiple words or morphs renders it a well-formed sequence.
    See \S\ref{ss:bh} for further discussion.}

@@

\noindent
However, there is no reason to regard these adaptations as a product of the synchronic phonology: no Yokuts root begins with a consonant, and there is no way to derive an initial CC cluster.
Furthermore, as \citet{Gamble1989} notes, there is no reason to believe that the adaptations actually occurred within Yokuts at all, since there was at best quite limited contact between Spanish and Yokuts speakers; the adaptations may have occurred in another language altogether.

@@

Finally, there are many cases where putative phonotactic restrictions are not extended to loanwords (e.g., \citealt{Clements1982}, \citealt[75]{Davidson1997}, \citealt{Fries1949}, \citealt{Holden1976}, \citealt{Ito1995a,Ito1995b}, \citealt[95]{Shibatani1973}, \citealt{Ussishkin2003}, \citealt{Vogt1954}; additional examples can be found throughout this dissertation).
For instance, native words in San Mateo Huave all end in a consonant, but 
\citet{Davidson1997} note that final unstressed syllables in Spanish loanwords are never repaired by epenthesis (e.g., \emph{verde} `green' $>$ [beɾde], *[beɾdej]).
Given the considerable disagreement about the nature of loanword adaptation at the present juncture, it may be premature to regard this inertness as strong evidence against the constraints in question, though it may be a useful diagnostic.

@@

Language games or speech disguises may also provide evidence for phonotactic knowledge \citep[e.g.,][]{Vaux2011}, assuming that these ``alternative phonologies'' are not qualitatively different from naturally-occurring language processes \citep[e.g.,][697]{Bagemihl1995}.
An example of language game evidence bearing on phonotactic questions can be found in \S\ref{ss:bh}, where a language game is used to argue for the hypothesis that root-internal vowel sequences in Turkish are subject to vowel harmony.

@@

Early generative phonologists posited that phonotactic ill-formedness derives solely from \emph{morpheme structure constraints}, restrictions on underlying representations \citep{Chomsky1965,SPE,SPR,Halle1962}.
\citet{Stanley1967} distinguishes between two types.
\emph{Segment structure constraints} impose restrictions on the underlying segment inventory.
For example, voicing is non-contrastive for /ts, tʃ, x/ in Russian \cite[22]{SPR}: [dz, dʒ, ɣ] appear in surface, but not underlying, representations.
This dissertation will have little to say about the nature of segment structure constraints.
Of more interest here are \emph{Sequence structure constraints} apply to underlying sequences; an example is given below.

@@

\noindent
This sequence structure constraint specifies that a consonant immediately after a word-initial stop is a liquid; this would preclude underlying */bnɪk/, for example.

However, \citet{Shibatani1973} shows quite convincingly that not all wordlikeness contrasts can be expressed as constraints on URs.
In German, for instance, obstruent voicing is contrastive, but neutralizes finally: e.g., [ɡʀaːt]-[ɡʀaːtə] `ridge(s)' vs.~[ɡʀaːt]-[ɡʀaːdə] `degree(s)'.
By hypothesis, the latter root ends in /d/, so the constraint against final voiced obstruents is specific to surface representations.
\citeauthor{Shibatani1973} claims, however, that German speakers judge voiced obstruent-final nonce words to be ill-formed.\footnote{
    Furthermore, voicing of final obstruents is usually lost in German loanword adaptation: e.g., English \emph{hot dog} becomes [hɑt dɔk] \citep[506]{Ussishkin2003}.}

>>>>

Phonotactic knowledge is evidenced in quite different ways than phonological knowledge.
In Russian, for instance, a process of anticipatory assimilation ensures that derived clusters of obstruents agree in voice.\footnote{
    For sake of discussion, the complex behavior of [v] has been ignored here.}

@@

\noindent
\citeauthor{PE} go on to label constraints on underlying representations as ``extensional'', and thus irrelevant to generative grammar, in contrast to ``intensional'' statements like phonological rules.
Regarding the facts about Russian obstruents discussed so far, it is clear  that the alternation facts are in some sense more privileged.
Were it the case that Russian speakers had not internalized a process of obstruent voicing assimilation, the only alternative explanation for the forms in (\ref{rovs}) is a massive system of phonologically conditioned suppletive allomorphy.
In contrast, were there no constraint on underlying obstruent voicing, however, it is not immediately obvious that anything would be different.
Presumably, this is what \citeauthor{PE} mean when they refer to constraints on underlying representations as ``computationally inert''.
There are also many cases where restrictions on underlying representations are specific to underlying representations and do not apply to surface representations.
For instance, hiatus is exceptionally rare in native Turkish roots; of the handful of examples in the Turkish Electronic Living Lexicon \citep{TELL}, many appear to be compounds (e.g., \emph{ısıalan} `endothermic', cf.~\emph{ısı} `heat', \emph{alan} `taker').
Yet numerous phonological processes give rise to hiatus in derived environments \citep[e.g.,][]{Kabak2007b}.
\citet{Silverman2000} observes that operations like reduplication and truncation also tend to introduce violations of restrictions that hold of underlying representations.
However, there are many facts which strongly suggest that speakers internalize phonotactic constraints.

@@

Speakers have difficulty producing \citep[e.g.,][]{Davidson2006a,Davidson2010,Rose2007,Vitevitch1997} and perceiving \citep{Dupoux1999,Kabak2007a,Massaro1983} certain types of nonce words judged to be phonotactically illicit.
For instance, \citet{GallagherInPress} finds that speakers of Quechua have difficulty repeating nonce words with multiple ejectives (e.g., [k'ap'i]), which do not occur in the language; in other words, multiple ejective sequences are not merely absent, but also difficult for these speakers to perceive or produce.

In English, sequences of adjacent obstruents which do not also agree in voice (e.g., \emph{a}[b.s]\emph{inth}) are quite rare within a word, and therefore a hetero-voiced obstruent cluster is a clue to the presence of a word boundary in running speech.
Infants \citep[e.g.,][]{Mattys2001b} and adults \citep{McQueen1998b,Norris1997} are thought to use this heuristic for word recognition in experimental settings.\footnote{
    It is important to distinguish the computations involved in using this word segmentation heuristic from those implicated by other phonotactic behaviors.
    Whatever the locus of [bnɪk]'s ill-formedness, for example, no segmentation into multiple words or morphs renders it a well-formed sequence.
    See \S\ref{ss:bh} for further discussion.}

@@

\noindent
However, there is no reason to regard these adaptations as a product of the synchronic phonology: no Yokuts root begins with a consonant, and there is no way to derive an initial CC cluster.
Furthermore, as \citet{Gamble1989} notes, there is no reason to believe that the adaptations actually occurred within Yokuts at all, since there was at best quite limited contact between Spanish and Yokuts speakers; the adaptations may have occurred to conform to the phonotactic restrictions of another language, also in contact with Yokuts.

@@

Finally, there are many cases where putative phonotactic restrictions are not extended to loanwords (e.g., \citealt{Clements1982}, \citealt[75]{Davidson1997}, \citealt{Fries1949}, \citealt{Holden1976}, \citealt{Ito1995a,Ito1995b}, \citealt[95]{Shibatani1973}, \citealt{Ussishkin2003}, \citealt{Vogt1954}; additional examples can be found throughout this dissertation).
For instance, native words in San Mateo Huave all end in a consonant, but 
\citet{Davidson1997} note that final unstressed syllables in Spanish loanwords are never repaired by epenthesis (e.g., \emph{verde} `green' $>$ [beɾde], *[beɾdej]).
Given the considerable disagreement about the nature of loanword adaptation at the present juncture, it may be premature to regard this inertness as strong evidence against the constraints in question, though it may be ultimately be a useful diagnostic.

@@

Language games or speech disguises may also provide evidence for phonotactic knowledge \citep[e.g.,][]{Vaux2011}, assuming that these ``alternative phonologies'' are not qualitatively different from naturally-occurring language processes \citep[e.g.,][697]{Bagemihl1995}.
An example of language game evidence bearing on phonotactic questions can be found in \S\ref{ss:bh}, where a language game is used to argue that root-internal vowel sequences in Turkish are subject to vowel harmony.

@@

Early generative phonologists posited that phonotactic ill-formedness derives solely from \emph{morpheme structure constraints}, restrictions on underlying representations \citep{Chomsky1965,SPE,SPR,Halle1962}.
\citet{Stanley1967} distinguishes between two types.
\emph{Segment structure constraints} impose restrictions on the underlying segment inventory.
For example, voicing is non-contrastive for /ts, tʃ, x/ in Russian \cite[22]{SPR}: [dz, dʒ, ɣ] appear in surface, but not underlying, representations.
This dissertation will have little to say about the nature of segment structure constraints.
Of more interest here are \emph{sequence structure constraints}, restrictions on underlying phoneme sequences. 
An example is given below.

@@

\noindent
This sequence structure constraint specifies that a consonant immediately after a word-initial stop is a liquid; this would preclude underlying */bnɪk/, for example (though it would erroneously rule out attested *[sn], *[sm] onsets).

However, \citet{Shibatani1973} argues that not all wordlikeness contrasts can be expressed as constraints on URs.
In German, for instance, obstruent voicing is contrastive, but neutralizes finally: e.g., [ɡʀaːt]-[ɡʀaːtə] `ridge(s)' vs.~[ɡʀaːt]-[ɡʀaːdə] `degree(s)'.
By hypothesis, the latter root ends in /d/, so the constraint against final voiced obstruents is specific to surface representations.
\citeauthor{Shibatani1973} claims, however, that German speakers judge voiced obstruent-final nonce words to be ill-formed.\footnote{
    Furthermore, voicing of final obstruents is usually lost in German loanword adaptation: e.g., English \emph{hot dog} becomes [hɑt dɔk] \citep[506]{Ussishkin2003}.}

### 2013-01-31 2.tex
Looks like this is wrapped up...!

\begin{figure}[t]
\centering
\includegraphics{agg.pdf}
\caption{Subjects freely use intermediate ratings when asked to rate how representative even and odd numbers were of ``even'' and ``odd'', respectively (\citealp{Armstrong1983}; ``1'' indicates ``most representative'')}
\label{agg}
\end{figure}

@@

\noindent
The logic of this implication is indisputable.
However, there is little empirical support for the claims that absolute and gradient well-formedness derive from similar principles; indeed, there have been no prior attempts to evaluate categorical and gradient models of wordlikeness on an equal footing.
In light of the complexities of gradient models, such an evaluation requires strong quantitative evidence for the superiority of gradient grammatical models.
The evaluation below represents a first attempt to fill this gap.
%FIXME clements and keyser

It is not that categorical models have been ignored by the literature on wordlikeness modeling, but rather that they have not been compared.
\citet{Frisch2000} and \citet{Vitevitch1997} find that speakers' wordlikeness ratings of multisyllabic words are correlated with a probabilistic measure of the well-formedness of the constituent syllables. 
Unfortunately, no attempt is made to control for the well-formedness of syllable contact clusters in these words: some of the stimuli have medial consonant clusters containing both voiced and voiceless obstruents (e.g., [ɡaɪbsaɪk]), something which is exceptionally rare in English simplex words (see \S\ref{ova}).
Similarly, \citet{Hayes2008a}, who compare their gradient model of wordlikeness against a set of English phonotactic constraints proposed by \citet{Clements1983}, first transform these constraints, several of which are without exception, into probabilities.
While this is consistent with their claim, that ``the ability to model gradient intuitions [is] an important criterion for evaluating phonotactic models'' \citep[382]{Hayes2008a}, little insight can be gained by annotating an exceptionless rule ``$p = 1.0$''. 
\citeauthor{Hayes2008a}'s principle precludes any attempt to test the hypotheses that underlies it, and therefore must be rejected.

% FIXME make this work: 
%(see \citealt{Sproat1993} and \citealt{Wrench2003} for articulatory investigations).

>>>>

\begin{figure}[t]
\centering
\includegraphics{agg.pdf}
\caption{When asked to rate how representative even and odd numbers were of ``even'' and ``odd'', respectively, subjects use intermediate ratings (from \citealp{Armstrong1983}; ``1'' indicates ``most representative'')}
\label{agg}
\end{figure}

@@

\noindent
The logic of this implication is indisputable.
However, there is little empirical support for the claims that absolute and gradient well-formedness derive from similar principles; indeed, there have been no prior attempts to evaluate categorical and gradient models of wordlikeness on an equal footing.
In light of the complexities of gradient models, such an evaluation requires strong quantitative evidence for the superiority of gradient grammatical models.
The evaluation below represents a first attempt to fill this gap.

It is not that categorical models have been ignored by the literature on wordlikeness modeling, but rather that they have not been compared.
\citet{Frisch2000} and \citet{Vitevitch1997} find that speakers' wordlikeness ratings of multisyllabic words are correlated with a probabilistic measure of the well-formedness of the constituent syllables. 
Unfortunately, no attempt is made to control for the well-formedness of syllable contact clusters in these words: some of the stimuli have medial consonant clusters containing both voiced and voiceless obstruents (e.g., [ɡaɪbsaɪk]), something which is exceptionally rare in English simplex words
%\citep{GormanInPressb}.
(see \S\ref{ova}). %FIXME
Similarly, \citet{Hayes2008a}, who compare their gradient model of wordlikeness against a set of English phonotactic constraints proposed by \citet{Clements1983}, first transform these constraints, several of which are without exception, into probabilities.
While this is consistent with their claim, that ``the ability to model gradient intuitions [is] an important criterion for evaluating phonotactic models'' \citep[382]{Hayes2008a}, little insight can be gained by annotating an exceptionless rule ``$p = 1.0$''. 
\citeauthor{Hayes2008a}'s principle precludes any attempt to test the hypotheses that underlies it, and therefore must be rejected.

### 2013-01-31 4.tex
Looks like this is wrapped up...!

Filtering the CELEX data according to the above criteria results in a list of 6,619 simplex words.
The full set of clusters and their frequencies are listed in Appendix \ref{clusters}.
The CELEX transcriptions of these words are then syllabified and phonologized using the same procedure as used in Chapter \ref{gradience} and described in Appendix \ref{syllabification}. 
In all, the sample contains 23 different medial coda and 40 different medial onsets.
Of the 920 ($= 21 \times 40$) medial clusters that would result from free combination of medial coda and medial onset, 174 (19\%) are attested.

@@

Since this model has numerous experimenter-defined parameters, a close replication of \citeauthor{Hayes2008a}'s original study is attempted: both their implementation and phonological feature specifications are used here, just as in Chapter \ref{gradience}.
Following \citet{HayesInPress}, dictionary entries are syllabified using the procedure described in Appendix \ref{syllabification}, and a novel feature [$\pm$ \textsc{Coda}] is added to allow the model to distinguish coda and onset consonants.
Also, following \citeauthor{Hayes2008a}, constraints are limited to those s
panning as many as three segments and the suggested ``accuracy schedule'' is used.
Since the maximum entropy model produces slightly different scores on each run, the worst-performing of 10 runs is reported here, following \citeauthor{Hayes2008a}.
This model has the poorest recall of any model; compared to the derived constraints baseline, the constraints induced by the maximum entropy model are narrower.
This is particularly clear regarding possible clusters with a nasal coda followed by a non-homorganic obstruent, like *[m.kl]: the vast majority of such clusters, which would be neutralized by \textsc{Nasal Place Assimilation}, are unattested, but many are erroneously assigned the highest possible score by the maximum entropy model.

>>>>

Filtering the CELEX data according to the above criteria results in a list of 6,619 simplex words.
The full set of clusters and their frequencies are listed in Appendix \ref{clusters}.
The CELEX transcriptions of these words are then syllabified and phonologized using a procedure
%the same procedure as used in Chapter \ref{gradience} and
described in Appendix \ref{syllabification}. 
In all, the sample contains 23 different medial coda and 40 different medial onsets.
Of the 920 ($= 21 \times 40$) medial clusters that would result from free combination of medial coda and medial onset, 174 (19\%) are attested.

@@

Since this model has numerous experimenter-defined parameters, a close replication of \citeauthor{Hayes2008a}'s original study is attempted: both their implementation and phonological feature specifications are used here.
%just as in Chapter \ref{gradience}. FIXME
Following \citet{HayesInPress}, dictionary entries are syllabified using the procedure described in Appendix \ref{syllabification}, and a novel feature [$\pm$ \textsc{Coda}] is added to allow the model to distinguish coda and onset consonants.
Also, following \citeauthor{Hayes2008a}, constraints are limited to those s
panning as many as three segments and the suggested ``accuracy schedule'' is used.
Since the maximum entropy model produces slightly different scores on each run, the worst-performing of 10 runs is reported here, following \citeauthor{Hayes2008a}.
This model has the poorest recall of any model; compared to the derived constraints baseline, the constraints induced by the maximum entropy model are narrower.
This is particularly clear regarding possible clusters with a nasal coda followed by a non-homorganic obstruent, like *[m.kl]: the vast majority of such clusters, which would be neutralized by \textsc{Nasal Place Assimilation}, are unattested, but many are erroneously assigned the highest possible score by the maximum entropy model.

### 2013-01-31 ack.tex
Looks like this is wrapped up...!

While many of my fellow students at Penn have had a hand in this document, none so much as my cohort: Toni Cook and Laurel MaKenzie. Without Joe Fruehwald, Marielle Lerner, Constantine Lignos, Brittany McLaughlin, and Hilary Prichard, it wouldn't have been nearly as much fun. Carly Fisher and Katherine McEldoon have provided me with unflagging support from afar. John Murdock and Hilary Prichard proofread the entire document numerous times.

@@

The ideas I first encountered as an undergraduate at the University of Illinois under the tutelage of Jennifer Cole and Richard Sproat, continue to guide my research. I am grateful to the members of the committee, Gene Buckley, Mark Liberman, and Rolf Noyer, for creating a nurturing environment where quantitative and theoretical approaches can meet. Insofar as I have acknowledged and understood the arguments of my predecessors, it is thanks to Steve Anderson, who graciously agreed to serve on my committee and to show me the wonders of New Haven. 
I do not expect to ever again have such a fearless advocate in the incomporable Charles Yang; the scientist he is is the one I strive to be.

While they were not involved in the preparation of this dissertation, it is hard to imagine that it would exist without Lila Gleitman, Morris Halle, and Bill Labov.

>>>>

While many of my fellow students at Penn have had a hand in this document, none so much as my cohort: Toni Cook and Laurel MaKenzie. 
Without Joe Fruehwald, Marielle Lerner, Constantine Lignos, Brittany McLaughlin, and Hilary Prichard, it wouldn't have been nearly as much fun. 
Carly Fisher and Katherine McEldoon have provided me with unflagging support from afar. 
John Murdock and Hilary Prichard graciously proofread the entire document numerous times.

@@

The ideas I first encountered as an undergraduate at the University of Illinois under the tutelage of Jennifer Cole and Richard Sproat continue to guide my research. 
I am grateful to the members of the committee, Gene Buckley, Mark Liberman, and Rolf Noyer, for creating a nurturing environment where quantitative and theoretical approaches can meet. 
Insofar as I have acknowledged and understood the arguments of my predecessors, it is thanks to Steve Anderson, who graciously agreed to serve on my committee and showed me the wonders of New Haven. 
I do not expect to ever again have such a fearless advocate in the incomporable Charles Yang; the scientist he is is the one I strive to be.
While they were not involved in the preparation of this dissertation, it is hard to imagine that it would exist without Lila Gleitman, Morris Halle, and Bill Labov.

### 2013-01-31 ms2.tex
Brought 4 up to speed with CLS 47 paper changes.
Looks like this is wrapped up...!

% fonts
\usepackage{mathspec}
\setmainfont[Mapping=tex-text]{Linux Libertine}
\setmathfont(Digits,Greek,Latin){Linux Libertine}

@@

\title{Categorical and gradient aspects of wordlikeness\thanks{Thanks to Gene Buckley, Charles Yang, and audiences at NELS 43 for helpful comments.}} 
\author{Kyle Gorman \\ University of Pennsylvania}
\date{November 2012 (comments welcome)}

@@

\abstract{
Gradient wordlikeness judgements do not necessarily imply that there is a gradient well-formedness system underlying them; gradient judgements may be an artifact of gradient rating tasks.
Dubious architectural assumptions are needed for speakers to report gradient well-formedness judgements.
Simple baselines better account for gradient well-formedness judgements than state-of-the-art computational models of gradient phonotactic knowledge.
}
\section{Introduction}
\input{2}
\appendix
\renewcommand{\arraystretch}{0.25}

\section{English wordlikeness ratings}
\label{ratings}

@@

This appendix describes an automated procedure used to process the CMU pronunciation dictionary entries, separating medial clusters from their flanking nuclei, parsing the resulting sequences into coda and onset, and reversing allophonic processes targeting medial clusters.

@@

\noindent
Consequently, pre-consonantal \emph{r} is assigned to the preceding nucleus.

@@

In English, when a medial consonant cluster is preceded by a stressed lax vowel, as \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel (\citealp[3]{Hammond1997}, \citealp{Treiman1990}). As \citet[55]{Harris1994} notes, however, when the medial cluster is also a valid onset, as in \emph{whi}[s.p]\emph{er}, \emph{ve}[s.ti]\emph{ge}, and \emph{mu}[s.k]\emph{et}, onset maximization will incorrectly assign the entire cluster to the onset and leave the lax vowel unchecked. For this reason, onset maximization parses are modified to assign the first consonant of a complex medial consonant cluster to the coda before a stressed lax vowel \citep[48]{Pulgram1970}.

\subsection{Phonologization}

The traditional analysis of affricates as single segments (e.g., \emph{SPE}:321f., \citealp[24]{Jakobson1961}) rather than stop-fricative sequences \citep[e.g.,][]{Hualde1988,Lombardi1990} is assumed here. In many languages, affricates pattern with simple onsets; for instance, Classical Nahua bans true onset clusters but permits the affricate series [ts, tʃ, tɬ] \citep[9]{Launey2011}. In English, however, affricates do not form complex onsets. Yet the stop and release phase of affricates cannot be separated by a syllable boundary, as predicted from the assumption they are single phonological units.

In English, [ŋ] has been analyzed as a pure allophone of /n/ before underlying /k, ɡ/ (with later deletion of /ɡ/ in some contexts; \citealt[65f.]{Borowsky1986}, \emph{SPE}:85, \citealt[62]{Halle1985a}), or as a phoneme in its own right \citep[e.g.,][]{Jusczyk2002,Sapir1925}. Onset [ŋ] is totally absent in onset position, where it cannot be followed by a /k, ɡ/ needed to derive the velar allophone, a fact predicted only by the former account, and English speakers have considerable difficulty producing initial [ŋ] \citep{Rusaw2009}. The allophonic analysis is assumed here. When followed by /k, ɡ/, [ŋ] is mapped to /n/. When not followed by a velar stop (i.e., finally), [ŋ] is analyzed as underlying /nɡ/.

%\section{Exemplification}
% 
%In (\ref{exemplification}), a few segmentations are given so as to demonstrate the principles the principles outlined above.
% 
%\begin{example}[Some examples of medial cluster segmentation]
%\label{exemplification}
%\begin{tabular}{l l l}
%a. & \emph{tranquil} & /\ldots{}n.kw\ldots{}/ & (coda nasal place assimilation, [Cw] onsets) \\
%b. & \emph{
%c. & \emph{
%d. & \emph{
%e. & \emph{
%f. & \emph{
%g. & \emph{
%h. & \emph{
%i. & \emph{
%j. & \emph{
%\end{tabular}
%\end{example}

>>>>

% fonts
\usepackage{mathspec}
\setmainfont[Mapping=tex-text]{Linux Libertine}
\setmathfont(Digits,Greek,Latin){Linux Libertine}
%\setmainfont[Mapping=tex-text]{Times New Roman}
%\setmathfont(Digits,Greek,Latin){Times New Roman}

@@

\title{Categorical and gradient aspects of wordlikeness\thanks{Thanks to Gene Buckley, Constantine Lignos, Hilary Prichard, Charles Yang, and audiences at NELS 43 for helpful comments.}}
\author{Kyle Gorman \\ University of Pennsylvania}
\date{Revised January 2013 (comments welcome)}

@@

\abstract{
Gradient wordlikeness judgements do not necessarily imply that there is a gradient well-formedness system underlying them; gradient judgements may be an artifact of gradient rating tasks.
Dubious architectural assumptions are needed for speakers to report gradient well-formedness judgements.
Simple baselines better account for gradient well-formedness judgements than state-of-the-art computational models of gradient phonotactic knowledge.
}
\section{Introduction}
\input{2}

\appendix
\renewcommand{\arraystretch}{0.25}
\section{English wordlikeness ratings}
\label{ratings}

@@

For every entry in the CELEX database, there is a corresponding broad syllabified transcription of the word in a Received Pronunciation accent. This appendix describes an automated procedure used to process these transcripts and to separate medial clusters from their flanking nuclei, parsing the resulting sequences into coda and onset, and reversing allophonic processes targeting medial clusters.

While the segmental content of these transcriptions is precise, the CELEX syllabifications are unsystematic. Given the absence of contrastive syllabification in English (if not all languages: see \citealp[221]{Blevins1995}, \citealp{Elfner2006}), any sequence of a medial consonant cluster and its flanking nuclei should receive the same syllabification in all words in which it occurs. This is not always the case with the CELEX transcriptions, however. For instance, the sequence [ɪstɹɪ] receives a different parse in \emph{chemistry} [ˈkɛ.mɪ.stɹɪ] and \emph{ministry} [ˈmɪ.nɪs.tɹɪ].\footnote{Note that word-final \emph{y} is usually lax in Received Pronunciation \citep[][II.294]{AOE}.} Consequently, these syllabifications are not used here.

@@

\noindent
Following \citet{Pierrehumbert1994}, pre-consonantal \emph{r} is assigned to the preceding nucleus.

@@

In English, when a medial consonant cluster is preceded by a stressed lax vowel, as \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel (\citealt[3]{Hammond1997}, \citealt{Treiman1990}). As \citet[55]{Harris1994} notes, however, when the medial cluster is also a valid onset, as in \emph{whi}[s.p]\emph{er}, \emph{ve}[s.ti]\emph{ge}, and \emph{mu}[s.k]\emph{et}, onset maximization will incorrectly assign the entire cluster to the onset and leave the lax vowel unchecked. For this reason, onset maximization parses are modified to assign the first consonant of a complex medial consonant cluster to the coda before a stressed lax vowel \citep[48]{Pulgram1970}.

\subsection{Phonologization}
\label{s:phonologization}

Following \citet{Pierrehumbert1994}, the traditional analysis of affricates as single segment (e.g., \emph{SPE}:321f., \citealp[24]{Jakobson1961}) rather than sequences of a stop and fricative \citep[e.g.,][]{Hualde1988,Lombardi1990} is adopted here.
In many languages, affricates pattern with simple onsets; for instance, Classical Nahua bans true onset clusters but permits the affricate series [ts, tʃ, tɬ] \citep[9]{Launey2011}.
Other languages, such as Polish, distinguish affricates and stop-fricative sequences \citep{Brooks1965}, providing further evidence that ``true'' affricates are represented as single segments (or single timing units), and in contrast with stop-fricative clusters \citep[34f.]{Clements1983}.

In English, [ŋ] has been analyzed as a pure allophone of /n/ before underlying /k, ɡ/ (with later deletion of /ɡ/ in some contexts; \citealt[65f.]{Borowsky1986}, \emph{SPE}:85, \citealt[62]{Halle1985a}), or as a phoneme in its own right \citep[e.g.,][]{Jusczyk2002,Sapir1925}. 
Onset [ŋ] is totally absent in onset position, where it cannot be followed by a /k, ɡ/ needed to derive the velar allophone, a fact predicted only by the former account, and English speakers have considerable difficulty producing initial [ŋ] \citep{Rusaw2009}. 
Following \citet{Pierrehumbert1994}, the allophonic analysis is assumed here. When followed by /k, ɡ/, [ŋ] is mapped to /n/. 
When not followed by a velar stop (i.e., finally), [ŋ] is analyzed as underlying /nɡ/.

% what about Fromkin 1973 on sprig time for hintler
